Leadership across cultures and contexts: an exploratory analysis

### 2020-05-16 leadership_across_cultures_contexts.Rmd
Minor edits in results.

Subsistence strategy was confounded with continental region. African text records were predominantly from horticultural societies, for instance, whereas North American text records were predominantly from hunter-gatherer societies (see Supplementary Figure \@ref(fig:regionsubsistence)).

@@

The vast majority of text records pertained to male leadership (`r male_leader_pct`%) and very few pertained to female leadership (`r female_leader_pct`%), with the remaining records not sex-specific. There were female leaders in almost all types of groups, however, except military and religious. Female leaders did tend to lead at the kin group and other residential subgroups, though: `r female_residential_pct`% of female leaders led at the residential subgroup level, for instance, compared to `r male_residential_pct`% of male leaders. The majority of text records which exclusively described female leaders were from horticulturalists (`r female_hort_pct`%). See Table S\@ref(tab:sexsubtable).

@@

We explored whether three features of our corpus of ethnographic texts -- date of publication, female gender of any author or co-author, and total pages of ethnography from each culture -- were associated with discovery of evidence for the `r length(all_study_vars)` dimensions in our study that coded for functions and qualities, benefits, and costs (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)). Dates of publication ranged from `r min(documents$d_publication_date)` to `r max(documents$d_publication_date)`, with a median year of `r median(documents$d_publication_date)` (see Supplementary Figure \@ref(fig:doc-years)). There were `r tab_author_gender['male']` male authors, `r tab_author_gender['female']` females authors, and `r tab_author_gender['unkown']` authors whose gender we could not determine (some authors were co-authors). Of the `r nrow(documents)` source documents, `r sum(documents$female_coauthor)` (`r signif(100*sum(documents$female_coauthor)/nrow(documents), 2)`%) had a female author or co-author. Total pages of ethnography on each culture in the PSF ranged from `r min(leader_cult[['ehraf_pages']])` to `r as.integer(max(leader_cult[['ehraf_pages']]))`, with a median of `r median(leader_cult[['ehraf_pages']])`.

@@

We fit `r length(all_study_vars)` logistic mixed effects models with each binary study dimension as the outcome; fixed effects for publication date, any female authors (True/False), and total pages of ethnography from each culture; and random intercepts for ethnographers nested within cultures. We then corrected for multiple comparisons by setting a false discovery rate of 5% [@benjamini1995]. After correction, none of the coefficients of female authorship or total pages of ethnography were significant, but the publication date coefficient was significant in nine models. More recent publications were more likely to report evidence for *Oratory skill*, *Group representative*, *Group determination/cohesiveness*, and *Knowledgeable/intelligent*, and less likely to report evidence for leaders imposing costs (*Punishment*, *Killer*) or receiving benefits (*Social services*, *Food*, and *Reduced risk of harm*). See Figure S\@ref(fig:pubdate) and supplementary information for details.

@@

Most of the standard deviations of the random author and culture intercepts for each model were between 0 and 1 (on the scale of the linear predictor). See Figures \@ref(fig:FunctionRE) and \@ref(fig:QualityRE). In a few cases, one or the other was exactly 0, which can occur in models like ours that have many random effect levels (e.g., `r length(unique(all_data$author_ID))` authors) and a relatively small data set [@bates2015].

>>>>

Subsistence strategy was confounded with continental region. African text records were predominantly from horticultural societies, for instance, whereas North American text records were predominantly from hunter-gatherer societies (see Figure S\@ref(fig:regionsubsistence)).

@@

The majority of text records pertained to male leadership (`r male_leader_pct`%) and very few pertained to female leadership (`r female_leader_pct`%), with the remaining records not sex-specific. There were female leaders in almost all types of groups, however, except military and religious. Female leaders did tend to lead at the kin group and other residential subgroups, though: `r female_residential_pct`% of female leaders led at the residential subgroup level, for instance, compared to `r male_residential_pct`% of male leaders. The majority of text records which exclusively described female leaders were from horticulturalists (`r female_hort_pct`%). See Table S\@ref(tab:sexsubtable).

@@

We explored whether three features of our corpus of ethnographic texts -- date of publication, female gender of any author or co-author, and total pages of ethnography from each culture -- were associated with discovery of evidence for the `r length(all_study_vars)` functions, qualities, benefits, and costs dimensions (see Tables \@ref(tab:qualfuncvars) and \@ref(tab:costbenefitvars)). Dates of publication ranged from `r min(documents$d_publication_date)` to `r max(documents$d_publication_date)`, with a median year of `r median(documents$d_publication_date)` (see Figure S\@ref(fig:doc-years)). There were `r tab_author_gender['male']` male authors, `r tab_author_gender['female']` females authors, and `r tab_author_gender['unkown']` authors whose gender we could not determine (some authors were co-authors). Of the `r nrow(documents)` source documents, `r sum(documents$female_coauthor)` (`r signif(100*sum(documents$female_coauthor)/nrow(documents), 2)`%) had a female author or co-author. Total pages of ethnography on each culture in the PSF ranged from `r min(leader_cult[['ehraf_pages']])` to `r as.integer(max(leader_cult[['ehraf_pages']]))`, with a median of `r median(leader_cult[['ehraf_pages']])`.

@@

We fit `r length(all_study_vars)` logistic mixed effects models with each binary dimension as the outcome; fixed effects for publication date, any female authors (True/False), and total pages of ethnography from each culture; and random intercepts for ethnographers nested within cultures. We then corrected for multiple comparisons by setting a false discovery rate of 5% [@benjamini1995]. After correction, none of the coefficients of female authorship or total pages of ethnography were significant, but the publication date coefficient was significant in nine models. More recent publications were more likely to report evidence for *Oratory skill*, *Group representative*, *Group determination/cohesiveness*, and *Knowledgeable/intelligent*, and less likely to report evidence for leaders imposing costs (*Punishment*, *Killer*) or receiving benefits (*Social services*, *Food*, and *Reduced risk of harm*). See Figure S\@ref(fig:pubdate) and supplementary information for details.

@@

Most of the standard deviations of the random author and culture intercepts for each model were between 0 and 1 (on the scale of the linear predictor). See Figures S\@ref(fig:FunctionRE) and S\@ref(fig:QualityRE). In a few cases, one or the other was exactly 0, which can occur in models like ours that have many random effect levels (e.g., `r length(unique(all_data$author_ID))` authors) and a relatively small data set [@bates2015].
